---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r setup, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%"
)
```
# Weakly Associated Vectors (WAVE) Sampling
Spatial data are generally auto-correlated, meaning that if two units selected are close to each other, then it is likely that they share the same properties. For this reason, when sampling in the population it is often needed that the sample is well spread over space. A new method to draw a sample from a population with spatial coordinates is proposed. This method is called `wave` (weakly associated vectors) sampling. It uses the less correlated vector to a spatial weights matrix to update the inclusion probabilities vector into a sample. For more details see Raphaël Jauslin and Yves Tillé (2020) https://doi.org/10.1007/s13253-020-00407-1.
  
  
## Installation

### CRAN version

```
install.packages("WaveSampling")
```

### Latest version 

You can install the latest version of the package `WaveSampling` with the following command:

``` r
# install.packages("devtools")
devtools::install_github("Rjauslin/WaveSampling")
```

## Simple example

This basic example shows you how to solve a common problem. Spatial coordinates from the function `runif()` are firstly generated. 

```{r}
library(WaveSampling)

N <- 144
n <- 48
X <- cbind(runif(N),runif(N))
head(X,10)
```

Inclusion probabilities `pik` is set up all equal with the function `rep()`.

```{r}
pik <- rep(n/N,times = N)
```


It only remains to use the function `wave()`,

```{r}
s <- wave(X,pik)
```

We can also generate a plot to observe the result.

```{r, fig.show='hold',fig.align='center',out.width='75%'}
library(ggplot2)
ggplot() +
  geom_point(data = data.frame(x = X[,1],y = X[,2]),
             aes(x = x,y = y),
             shape = 1,
             alpha = 0.2)+
  geom_point(data = data.frame(x = X[s == 1,1],y = X[s == 1,2]),
             aes(x,y),
             shape = 16,
             colour = "black")+
  theme_void()

```
