---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r setup, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%"
)
```
# Weakly associated vectors (wave)
Spatial data are generally autocorrelated, meaning that if two units selected are close to each other, then it is likely that they share the same properties. For this reason, when sampling in the population it is often needed that the sample is well spread over space. A new method to draw a sample from a population with spatial coordinates is proposed. This method is called `wave` (weakly associated vectors) sampling. It uses the less correlated vector to a spatial weights matrix to update the inclusion probabilities vector into a sample.
  
## Build 
[![Build Status](https://travis-ci.org/RJauslin/WaveSampling.svg?branch=master)](https://travis-ci.org/RJauslin/WaveSampling)

## Installation

You can install the released version of the package `WaveSampling` with the following command

``` r
# install.packages("devtools")
devtools::install_github("Rjauslin/wave")
```

## Simple example

This basic example shows you how to solve a common problem. Spatial coordinates from the function `runif()` are firstly generated. 

```{r}
library(WaveSampling)

N <- 144
n <- 48
X <- cbind(runif(N),runif(N))
head(X,10)
```

Inclusion probabilities `pik` is set up all equal with the function `rep()`.

```{r}
pik <- rep(n/N,times = N)
```


It only remains to use the function `wave()`,

```{r}
s <- wave(X,pik)
```

We can also generate a plot to see

```{r, fig.show='hold',fig.align='center',out.width='75%'}
library(ggplot2)
ggplot() +
  geom_point(data = data.frame(x = X[,1],y = X[,2]),
             aes(x = x,y = y),
             shape = 1,
             alpha = 0.2)+
  geom_point(data = data.frame(x = X[s == 1,1],y = X[s == 1,2]),
             aes(x,y),
             shape = 16,
             colour = "black")+
  theme_void()

```


## Performance

As explained on the website of Microsoft R open https://mran.microsoft.com/open, R is designed to use only one thread but could be linked with a multi-threaded version of BLAS/LAPACK. The package is implemented with the package RcppArmadillo http://arma.sourceforge.net/ that provide an integration of various matrix decompositions with LAPACK library. Intel MKL that is used by the Microsoft R open use a multi-threaded version of BLAS/LAPACK. Hence the package could gain time from the Microsoft R open. (Not tested, but linking R with OpenBLAS should also work).


```
N <- 500
n <- 100
X <- as.matrix(cbind(runif(N),runif(N)))
pik <- rep(n/N,N)
system.time(s <- wave(X,pik,tore = T,shift =T,comment = T))

```

#### R 3.6.1
```
user        system    elapsed 
114.52      1.90      118.39
```

#### MRO 3.5.3

```
user        system     elapsed 
96.03       1.38       25.82
```
