---
title: "Weakly Associated Vectors"
author: "Raphaël Jauslin"
date: "`r Sys.Date()`"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Introduction}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r setup, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
library(wave)
```


## Introduction
Spatial data are generally autocorrelated, meaning that if two units selected are close to each other then it is likely that they share the same properties. For this reason, when sampling in the population it is often needed that the sample is well spread over space. We propose here a new way of estimating a sample from a population with spatial coordinates. This method is called `wave` (weakly associated vectors). It uses the less correlated vector to a spatial weights matrix **W** to update the inclusion probabilities vectors into a sample **s**.


## 2-D sampling
We propose here two examples showing how to use the function `wave()`. The first one is an example where the inclusion probabilities of the units in the population are all equal and the spatial coordinates are on a regular grid. The second one is an example where the spatial coordinates are generated from an `runif()` distribution. 


### Regular grid

Firtly we need to generate the spatial coordinates. We then create an 10 x 10 regular grid as follows,

```{r}
N <- 144
n <- 48
x <- seq(1,sqrt(N),1)
X <- as.matrix(cbind(rep(x,times = sqrt(N)),rep(x,each = sqrt(N))))
head(X,10)
```

Now we need the inclusion probabilities `pik`. Hence, we generate a vector `pik_equal` that have all the same values and `pik_unequal` that have all values different but sum up to `n = 10`.

```{r}
pik_equal <- rep(n/N,times = N)
pik_unequal <- sampling::inclusionprobabilities(runif(N),n) 

head(pik_equal)
sum(pik_equal)

head(pik_unequal)
sum(pik_unequal)
```

Finally, we have the components to use `wave()`.

```{r}
s_equal <- wave(X,pik_equal,tore = TRUE,jitter = TRUE)
head(s_equal)
s_unequal <- wave(X,pik_unequal,tore = TRUE,jitter = TRUE)
head(s_unequal)

```

Graphically this gives,

```{r, fig.show='hold'}
library(ggplot2)
X <- data.frame(x = X[,1],y = X[,2])
ggplot() +
  geom_point(data = X,aes(x = x,y = y),pch = 1,alpha = 0.2)+
  geom_point(data = data.frame(x = X[s_equal == 1,1],y = X[s_equal == 1,2]),aes(x,y),colour = "black")+
  theme_light()+
  theme(panel.grid.major = element_blank(),
        panel.grid.minor = element_blank())
ggplot() +
  geom_point(data = X,aes(x = x,y = y),pch = 1,alpha = 0.2)+
  geom_point(data = data.frame(x = X[s_unequal == 1,1],y = X[s_unequal == 1,2]),aes(x,y),colour = "black")+
  theme_light()+
  theme(panel.grid.major = element_blank(),
        panel.grid.minor = element_blank())


```

### Random uniform


Firtly we need to generate the spatial coordinates. We then create randomly uniform coordinates from the function `runif()`,

```{r}
N <- 144
n <- 48
X <- cbind(runif(N),runif(N))
head(X,10)
```

We use the same inclusion probabilities as the previous example, we then just have to recalculate the samples.

```{r}
s_equal <- wave(X,pik_equal,tore = FALSE,jitter = FALSE)
head(s_equal)
s_unequal <- wave(X,pik_unequal,tore = FALSE,jitter = FALSE)
head(s_unequal)

```

Graphically this gives,

```{r, fig.show='hold'}
library(ggplot2)
X <- data.frame(x = X[,1],y = X[,2])
ggplot() +
  geom_point(data = X,aes(x = x,y = y),pch = 1,alpha = 0.2)+
  geom_point(data = data.frame(x = X[s_equal == 1,1],y = X[s_equal == 1,2]),aes(x,y),colour = "black")+
  theme_light()+
  theme(panel.grid.major = element_blank(),
        panel.grid.minor = element_blank())
ggplot() +
  geom_point(data = X,aes(x = x,y = y),pch = 1,alpha = 0.2)+
  geom_point(data = data.frame(x = X[s_unequal == 1,1],y = X[s_unequal == 1,2]),aes(x,y),colour = "black")+
  theme_light()+
  theme(panel.grid.major = element_blank(),
        panel.grid.minor = element_blank())


```



## 1-D sampling



## Performance







## Styles

The `html_vignette` template includes a basic CSS theme. To override this theme you can specify your own CSS in the document metadata as follows:

    output: 
      rmarkdown::html_vignette:
        css: mystyles.css


You can enable figure captions by `fig_caption: yes` in YAML:

    output:
      rmarkdown::html_vignette:
        fig_caption: yes

Then you can use the chunk option `fig.cap = "Your figure caption."` in **knitr**.


Also a quote using `>`:

> "He who gives up [code] safety for [code] speed deserves neither."
([via](https://twitter.com/hadleywickham/status/504368538874703872))
